Rapid, (semi-?)Supervised Generation of Novel Percussive Sounds With A Tractable Virtual Synthesizer

### 2020-01-03 paper.tex
section 3 + 4

\section{Introduction}
\subsection{Terminology}
\todo[inline,color=green!30]{I'd like to define terms such as VSTs, audio effects, Synthesizers, paramters and presets here. I throw these terms around a lot and I think some might be confused by their meaning. Is this the appropriate place for that? I'd like the read to know these definionts before reading the intro. I can try to keep it brief and engaging}
\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops(percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects, which some producers may find daunting or time consuming to work with from scratch. In many cases, VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. Producers may further modify these presets until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers that a producer might find  interesting. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for promising avenues towards our motivational goal.\\

@@

\subsection{Related Work and Contradistinctions}
\label{related}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal \cite{nsynth2017}\colorbox{green!=40}{will add more references} and recently, Variational AutoEncoders (VAE's) have been utilized for generation of short percussive samples \cite{aouameur2019neural,ramires2019timbfeat}. In this work however, we opt to use digital signal processing methods to create a virtual synthesizer for the generation of audio signals, as it provides several unique advantages:
\begin{enumerate}[label=\roman*]
  \item Fast, offline rendering of audio with no relience on GPU, currently not possible with state of the art models such as parallel WaveGan \cite{yamamoto2019parallel} and parallell WaveNet \cite{oord2017parallel}. 
  \item Rendering at high sampling rates with little performance issues: Performance speed being a common issue, the standard sampling rate in most audio generation work utilizing neural networks appears to be under 24 khz \cite{yamamoto2019parallel,oord2017parallel,aouameur2019neural,ramires2019timbfeat}. However, a significant number of untrained human ears can detect a change in quality of audio between sampling rates of 192 khz and the industry standard of 44.1 khz \cite{reiss2016meta}, with a dramatic increase in quality detection after training. Therefore we can safely assume that most producers would prefer their audio samples have sampling rates of 44.1 khz or higher. 
  \item Neural networks are often viewed as unexplainable black box solutions. Some models such as VAE's can learn an underlying latent space of parameters and capture the "essence" of the different labels in a dataset. However, these spaces are learned in an unsupervised manner and must be manually analysed, perhaps extensively, before they can be understood \cite{esling2018generative}. The use of a virtual synthesizer for audio generation makes our parameters readily understandable and easily modifiable. \\

@@

\subsection{The Ear}
We defined an ear as any method of evaluating a piece of audio, capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation, what we require from the ear is to give us probabilities of an audio sample belonging to various sound (drum and none-drum) categories. As our synthesizer outputs are deterministic for all programs, allowing us to associate the categorical probabilities of each sound with the program that generated it; navigating our synthesis towards parameters which give us the desired sonic output.\\

%  The most basic representation of a piece of digital audio is via list of numbers and a sampling-rate. For simplicity, we fix the sampling rate to 41,000 per second across all components of this project. An array representation of audio is the representation of the audio within the time domain. More commonly, the frequency domain is used for MIR tasks.
We have implemented several methods of feature extraction and successfully separated different drum categories in interesting ways. 
\todo[inline, color=green!30]{is this the right place to show off some t-sne graphs of how various feature extraction methods seperated drum categories?}
A common pitfall of most feature extraction methods implemented is that they do not consider the temporal trends within the audio, making them highly susceptible to errors when evaluating audio that should not belong to any of the categories.\\
Analysis of the frequency content of a piece of audio using windowing functions is a common MIR procedure. Care must be taking such that information critical to percussion categorization such as envelope and temporal changes are not lost in this process. For example, we do not want a reversed kick sample to be labeled as a kick, or even a drum, even if it has the correct frequency content. There are two interconnected temporal trends to consider: 
\begin{itemize}
    \item Change in amplitude within each frequency bin. Harder to measure and utilize and possibly not as important for our purposes, considering the short length of our samples.
    \item Overall amplitude change within a sample. Closely linked to ADSR\\. 
\end{itemize}{}
Temporal trends may be deduced from a spectrum representation of sound. A convolution net trained on our drum dataset, combined with an algorithm ADSR extraction algorithm led to an ear which can reliably reject junk data and samples with inappropriate wave-shapes. \todo[inline,color=green!30]{This paragraph is lying, I haven't done the ADSR algorithm part yet. Highest priority after writing this paper.}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{images/cnn_ear.png}
\caption{Model working well on a test set}
\label{fig:ok ear}
\end{figure}
\todo[inline,color=green!30]{ will add f-measure calculations and other confidence inducing graphs (for example I have a grid of spectrums with the ear's guess vs the actual label under each spectrum, I can colorize the grid depending on whether the guess was correct or not). Code in the notebooks. After updating ear, I will make them look presentable and add them.}
%
% How Can The Synth and Ear Interact?
%
\subsection{How Can The Synth and Ear Interact?}

@@

With our Synth and Ear in place, we can quickly generate programs for our Synth, create the corresponding audio sample and get the scores for the audio sample. If the ear returns a vector of categorical probabilities, we can assign categorical ranks to the audio sample. For example, the category with the highest probability will have the rank of 1, second highest would rank 2nd, and the lowest probability will have the rank of n (assuming n categories).\textbf{
But how can we use this to create percussion sounds?}\\\\
We utilize our tools in three different methods of generation, with their specific imlpementation discussed in section \ref{gens}.
\begin {enumerate} [label=(\roman*)]
\item Random Generation: We randomly create a drum programs until a sound belonging to our category is produced. We hypothesize that due to the speed of our experiments, it is a viable approach.
\item Multivariate Distribution Sampling: We attempt to learn the joint distribution of our parameters in order to maximize generation of a certain drum category. 
\item Genetic Search: We treat our parameters as genes and attempt to use evolutionary search algorithms to find programs that the ear deems most fit.\\
\end {enumerate} 
\todo[inline,color=blue!30]{As seen in Figure \ref{fig:rank portions}, there is a bias in random generation towards making higher ranking hats and kicks and lower ranking shakers, claps. Is this bias because of Synth structure or a problem with the ear? Will analyse further after updating ear}
\begin{figure}[h!]
\centering
\includegraphics[width=1\linewidth]{images/random_ranks.png}
\caption{\colorbox{green!=40}{will update and make labels bigger} Showing proportions of categorical ranks for around 50,000 randomly generated sounds. Only showing ranks 1-5. About 90\% received the rank of 1 (blue portion of the bar) for the "synth" category, meaning that the ear correctly gave the highest probability to the sound being from a synth combination. We are interested in the small percentages of random generations that trick the ear due to coincidentally being similar to actual drums (i.e those who received a rank of 1) }
\todo[inline, color=blue!30]{When looking for the best sounds of each category, could use ranks 2 and 3 etc. as well, however:
\unexpanded{\unexpanded{
\begin{itemize}
    \item We must figure out how to penalize their parameters
    \item The task requires even more confidence in ear,
\end{itemize} 
}}}

\label{fig:rank portions}
\end{figure}
\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category.

\subsection{Methodologies}
\subsubsection{Random Generation}

\subsubsection{Multivariate Normal}
Some success and extremely fast and scalable once the parameter space is calculated. Currently our parameters are array indices mapping to different values for Synth components. This makes the assumption that our discrete parameter space is normally distributed or differentiable in-appropriate and prone to errors.

\subsubsection{Evolutionary Search}
Implemented evolutionary search using DEAP \citep{DEAP_JMLR2012}. A parameter set is assumed to be an individual with each parameter being its genes. To measure an individuals fitness, we use the formula $fitness=score-rank$. As generations evolve, a Hall-of-Fame list with a fixed size tracks the best individuals to ever live in any given generation, updating itself after each batch of off-springs is evaluated.\\
To ensure diversity of the HOF individuals, the fitness score should also consider some measure of uniqueness. A weighted minkowski distance \footnote{https://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.distance.minkowski.html} between the genes of an offspring and the HOF individuals may be a good measurement.\\
A generation of $N$ individuals are initialized(randomly or using already existing ranked parameters). To create the offspring generation, each generation goes through mutations and in-place gene crossovers. The new generation consists of the top 80\% of the offspring generation and 20\% new individuals to quicken gene diversity. If the new individuals are coming directly from previously measured high scoring parameter sets, we ensure that they cannot be directly added to HOF without going through mutations and crossovers first.

\bibliographystyle{abbrv}
    \bibliography{nime-references} 
\end{document}

>>>>

\section{Introduction}
\subsection{Terminology}
\todo[inline,color=green!30]{I'd like to define terms such as VSTs, audio effects, Synthesizers, paramters and presets here. I throw these terms around a lot and I think some might be confused by their meaning. Is this the appropriate place for that? I'd like the reader to know these definitions before reading the intro. I can try to keep it brief and engaging}
\subsection{Background and Motivation}
The rise of Digital Audio Workstations (DAW) \cite{leider2004digital} and Virtual Studio Technology (VST) based plug-ins \cite{tanev2013virtual} have rapidly transformed the sonic and material landscape of music production in the recent years. Coupled with this rise in popularity is a vast array of commercial products and services dedicated to satiating the need of amateur and professional music producers for unique sounds; most commonly via audio samples: one-shot drum samples, long sustained notes (commonly referred to as pads or textures), and loops(percussive or melodic) are common deliverables. Two notable examples of these commercial services are \textit{loopmasters}\footnote{loopmasters.com} and \textit{splice.com}\footnote{splice.com}. Furthermore, VST plug-ins can emulate complex audio synthesizers and effects, which some producers may find daunting or time consuming to work with from scratch. In many cases, VST plug-in vendors or unaffiliated enthusiasts sell additional presets for these plugins, targeted towards producers who do not have the time or interest in creating their own. Producers may further modify these presets until their desired sound is reached.\\
Our work is motivated by the idea of finding new, convenient methods for the expansion of a music producer's library of sounds. Primarily with generation of novel, one-shot audio samples but also with automated search and creation of presets for virtual synthesizers that a producer might find  interesting. Using the generation of short, percussive audio samples as a starting point, this project is a proof of concept for promising avenues towards our motivational goal.\\

@@

\subsection{Related Work and Contradistinctions}
\label{related}
Numerous deep, neural network models have been proposed and utilized for the purpose of signal generation in recent years. WaveGans and WaveNet have been subject to significant improvements and experiments since their proposal \colorbox{green!=40}{\cite{nsynth2017} + Many more references to be added.} Even more recently, Variational AutoEncoders (VAE's) have been utilized for generation of short percussive samples \cite{aouameur2019neural,ramires2019timbfeat}. In this work however, we opt to use digital signal processing methods to create a virtual synthesizer for the generation of audio signals, as it provides several unique advantages:
\begin{enumerate}[label=\roman*]
  \item Fast, offline rendering of audio with no relience on GPU, currently not possible with state of the art models such as parallel WaveGan \cite{yamamoto2019parallel} and parallell WaveNet \cite{oord2017parallel}. 
  \item Rendering at high sampling rates: Performance speed being a common issue, the standard sampling rate in most audio generation work utilizing neural networks appears to be under 24 khz \cite{yamamoto2019parallel,oord2017parallel,aouameur2019neural,ramires2019timbfeat}. However, a significant number of untrained human ears can detect a change in quality of audio between sampling rates of 192 khz and the industry standard of 44.1 khz \cite{reiss2016meta}, with a dramatic increase in quality detection after training. Therefore we can safely assume that most producers would prefer their audio samples have sampling rates of 44.1 khz or higher. 
  \item Neural networks are often viewed as unexplainable black box solutions. Some models such as VAE's can learn an underlying latent space of parameters and capture the "essence" of the different labels in a dataset. However, these spaces are learned in an unsuperv manner and must be manually analysed, perhaps extensively, before they can be understood \cite{esling2018generative}. The use of a virtual synthesizer for audio generation makes our parameters readily understandable and easily modifiable. \\

@@

\subsection{The Ear}
We defined an ear as any method of evaluating a piece of audio, capable of "listening" to a piece of audio and giving it a score (or a list of scores) based on how well it satisfies certain criteria. Since in this work we are mainly focused on percussive generation, what we require from the ear is to give us probabilities of an audio sample belonging to various sound (drum and none-drum) categories. As our synthesizer outputs are deterministic for all programs, allowing us to associate the categorical probabilities of each sound with the program that generated it; navigating our synthesis towards parameters which give us the desired sonic output.\\
%  The most basic representation of a piece of digital audio is via list of numbers and a sampling-rate. For simplicity, we fix the sampling rate to 41,000 per second across all components of this project. An array representation of audio is the representation of the audio within the time domain. More commonly, the frequency domain is used for MIR tasks.
We have implemented several methods of feature extraction and successfully separated different drum categories in interesting ways. 
\todo[inline, color=green!30]{is this the right place to show off some t-sne graphs of how various feature extraction methods seperated drum categories?}
A common pitfall of most feature extraction methods implemented is that they do not consider the temporal trends within the audio, making them highly susceptible to errors when evaluating audio that should not belong to any of the categories.\\
Analysis of the frequency content of a piece of audio using windowing functions is a common MIR procedure. Care must be taking such that information critical to percussion categorization such as envelope and temporal changes are not lost in this process. For example, we do not want a reversed kick sample to be labeled as a kick, or even a drum, even if it has the correct frequency content. There are two interconnected temporal trends to consider: 
\begin{itemize}
    \item Change in amplitude within each frequency bin. Harder to measure and utilize and possibly not as important for our purposes, considering the short length of our samples.
    \item Overall amplitude change within a sample. Closely linked to ADSR\\. 
\end{itemize}{}
Temporal trends may be deduced from a spectrum representation of sound. A convolution net trained on our drum dataset, combined with an algorithm ADSR extraction algorithm led to an ear which can reliably reject junk data and samples with inappropriate wave-shapes. \todo[inline,color=green!30]{This paragraph is lying, I haven't done the ADSR algorithm part yet. Highest priority after writing this paper.}
\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{images/cnn_ear.png}
\caption{Model working well on a test set}
\label{fig:ok ear}
\end{figure}
\todo[inline,color=green!30]{ will add f-measure calculations and other confidence inducing graphs (for example I have a grid of spectrums with the ear's guess vs the actual label under each spectrum, I can colorize the grid depending on whether the guess was correct or not). Code in the notebooks. After updating ear, I will make them look presentable and add them.}
%
% How Can The Synth and Ear Interact?
%
\begin{figure}[H]
\centering
\includegraphics[width=1\linewidth]{images/random_ranks.png}
\caption{\colorbox{green!=40}{labels bigger + show other stack sizes} Showing proportions of categorical ranks for around 50,000 randomly generated sounds. Only showing ranks 1-5. About 90\% received the rank of 1 (blue portion of the bar) for the "synth" category, meaning that the ear correctly gave the highest probability to the sound being from a synth combination. We are interested in the small percentages of random generations that trick the ear due to coincidentally being similar to actual drums (i.e those who received a rank of 1) }
\todo[inline, color=blue!30]{When looking for the best sounds of each category, could use ranks 2 and 3 etc. as well, however:
\unexpanded{\unexpanded{
\begin{itemize}
    \item We must figure out how to penalize their parameters
    \item The task requires even more confidence in ear,
\end{itemize} 
}}}

\label{fig:rank portions}
\end{figure}
\subsection{How Can The Synth and Ear Interact?}

@@

With our Synth and Ear in place, we can quickly generate programs for our Synth, create the corresponding audio sample and get the scores for the audio sample. If the ear returns a vector of categorical probabilities, we can assign categorical ranks to the audio sample. For example, the category with the highest probability will have the rank of 1, second highest would rank 2nd, and the lowest probability will have the rank of n (assuming n categories).\textbf{
But how can we use this to create percussion sounds?}\\\\
We utilize our tools in three different methods of generation, with their specific imlpementation discussed in section \ref{gens}.
\begin {enumerate} [label=(\roman*)]
\item Random Generation: We randomly create a drum programs until a sound belonging to our category is produced. We hypothesize that due to the speed of our experiments, it is a viable approach.
\item Multivariate Distribution Sampling: We attempt to learn the joint distribution of our parameters in order to maximize generation of a certain drum category. 
\item Genetic Search: We treat our parameters as genes and attempt to use evolutionary search algorithms to find programs that the ear deems most fit.\\
\end {enumerate} 
\todo[inline,color=blue!30]{As seen in Figure \ref{fig:rank portions}, there is a bias in random generation towards making higher ranking hats and kicks and lower ranking shakers, claps. Is this bias because of Synth structure or a problem with the ear? Will analyse further after updating ear}

\section{Novel Generations}
\label{gens}
\subsection{The Goal}
 Given a category, our aim is to use random generations to guide an algorithm towards novel generation of drums for that category. We evaluate each method by measuring how many samples of categories \textit{kick,snare} and \textit{hat} can be produced in under a minute. Finally we show results of a survey conducted to measure the quality of these samples. 
\subsection{General Methods}
\subsubsection{Random Generation}
To establish a benchmark, we seek to establish the likelihood of generating samples of each category if we generate programs for various stack sizes by random selection of parameters. \\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-a}
\includegraphics[width=0.48\linewidth]{example-image-a}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:rand-graphs}
\end{figure}
\subsubsection{Multivariate Normal}
Some success and extremely fast and scalable once the parameter space is calculated. A well known issue with this apporach is that the parameter space for virtual synthesizers are often neither differenciable nor gaussian \cite{esling2019universal} (a couple more papers also talked about this, I have to find them). Currently our parameters are array indices mapping to different values for Synth components. This makes the assumption that our discrete parameter space is normally distributed or differentiable in-appropriate and prone to errors. \\
\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-b}
\includegraphics[width=0.48\linewidth]{example-image-b}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:multi-graphs}
\end{figure}

\subsubsection{Evolutionary Search}
Implemented evolutionary search using DEAP \citep{DEAP_JMLR2012}. A parameter set is assumed to be an individual with each parameter being its genes. To measure an individuals fitness, we use the formula $fitness=score-rank$. As generations evolve, a Hall-of-Fame list with a fixed size tracks the best individuals to ever live in any given generation, updating itself after each batch of off-springs is evaluated.\\
To ensure diversity of the HOF individuals, the fitness score should also consider some measure of uniqueness. A weighted minkowski distance \footnote{https://docs.scipy.org/doc/scipy/reference/generated/scipy.spatial.distance.minkowski.html} between the genes of an offspring and the HOF individuals may be a good measurement.\\
A generation of $N$ individuals are initialized(randomly or using already existing ranked parameters). To create the offspring generation, each generation goes through mutations and in-place gene crossovers. The new generation consists of the top 80\% of the offspring generation and 20\% new individuals to quicken gene diversity. If the new individuals are coming directly from previously measured high scoring parameter sets, we ensure that they cannot be directly added to HOF without going through mutations and crossovers first.\\

\begin{figure}[H]
\centering
\begin{subfigure}[b]{\linewidth}
\includegraphics[width=0.48\linewidth]{example-image-c}
\includegraphics[width=0.48\linewidth]{example-image-c}
\label{fig:1stack}
\setcounter{subfigure}{2}%
\end{subfigure}
\caption{\colorbox{green!=40}{todo}Graphs showing the effectiveness of this method for producing chosesn categories in under a minute}
\label{fig:evo-graphs}
\end{figure}
\subsubsection{Survey of Generated Drums}
\label{survey}
\colorbox{green!=30}{Finalize ear,generate,make survey.} The survey will be sent out. I know a few producers who are willing to do this.
3 algorithms, 3 categories, 10 samples for each algorithm and category pair, 2 seconds to listen to each sample-> 3 minutes for listening
I imagine survey will be around 5-6 minutes long.
\subsubsection{Contrasting Ear Scores With Survey Results}
\colorbox{green!=30}{pending section \ref{survey}}
\section{Code and Database}
\colorbox{green!=30}{One last thing to do, will take a 1 day}\\
\colorbox{blue!=30}{Should this section be at the start of the paper?Appendix?}\\
talk about  categories, sources, dataprocessing. 
Code for the data processing is written, I need to download the no-copyright sounds, run the code and host the results. 
Give link for people to download it, show where to find the scripts for database expansion.
\section{Summary and Conclusions}
\colorbox{green!=30}{Points I want to make:}
\begin{itemize}
    \item Our methodology was successful in generating novel examples of some percussion categories.
    \item With the right implementation of DSP methods to create a synth, Genetic search is viable for finding  desired programs and parameters.
    \item When possible, using DSP methods instead of NNs is preferable in many ways(speed,computation,hardware requirements).\\
    \colorbox{green!=30}{Include some speed measurements in appendix?}
    \item ...
\end{itemize}
\section{Future Work}
\colorbox{blue!=30}{Should really think this through}
\begin{itemize}
    \item Potential for other generative methods (e.g reinforcement learning)
    \item Making the synth's sound range more diverse while maintaining speed and efficiency
\end{itemize}
\bibliographystyle{abbrv}
    \bibliography{nime-references} 
\end{document}
